Large Scale Marine Protected Areas

### 2019-01-24 docs/PIPA_Palau.tex
Updated some figures and tables that had some details. Updated manuscript to send to John

\begin{abstract}
Large-scale Marine Protected Areas (LSMPAs) have seen a significant
increase over the last years. Fishing effort is effectively eliminated
within these protected areas upon implementation. The benefits of
reducing effort have been largely studied, but little empirical works
evaluate how vessels react and redistribute after an MPA is created. The
economic and ecological implications of displacing fishing effort are
not yet fully understood. We use identification of fishing activity via
Automatic Identification Systems (AIS) and causal inference techniques
to provide the first analysis of behavioral changes and spatial
redistribution of tuna purse seiners due to the implementation of a
Large Scale Marine Protected Area in the Pacific Ocean. Our work
provides three main findings: 1) aggregate fishing effort remains
relatively unaffected; 2) vessels that fished inside the protected area
redistribute to adjacent waters; and 3) we observe a crowding effect for
the first months after implementation. Our results not only provide an
impact evaluation of the effect of LSMPAs on fishing activity, but
provide insights into vessel redistribution dynamics, which may have
ecological and economic implications for marine conservation and fisheries
management. As countries continue to implement LSMPAs as a way to
reach the stated 10\% target of ocean protection, managers should
consider how fishing effort will change in space and through time to
ensure that fishing effort is not just displaced elsewhere, leading to
overfishing in adjacent waters. While LSMPAs can provide a wide
range of benefits, their implementation must be accompained with
traditional fisheries management to maximize their effectiveness.

@@

\usepackage{pdflscape}
\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}

\begin{document}

@@

\section{Introduction}\label{introduction}
\dropcap{M}arine Protected Areas (MPAs) are intended to safeguard parts of the
ocean from fishing and other extractive activities. Current
international goals aim to protect 10\% of the ocean environment by
2020. In an effort to meet this target, there has been a rapid
increase recently in MPA coverage \citep{wood_2008,sala_2018}, largely driven by
a small number of Large Scale Marine Protected Areas (LSMPAs)\footnote{See, for example, \citep{game_2009,singleton_2014,boonzaier_2016,mccauley_2016,alger_2017}.}.
Today, a small number of LSMPAs make up at least 80\%
of the managed areas in the ocean \citep{toonen_2013}.

@@

\hypertarget{large-scale-marine-protected-areas}{%
\subsection{Large Scale Marine Protected
Areas}\label{large-scale-marine-protected-areas}}

@@

The amount of data gathered by GFW is dependent on the number of antennas and
satellites that can receive signals. The total satellite count increased from 3 to 6 on 
June 1\textsuperscript{st}
2014 , and then from 6 to 10 on January
1\textsuperscript{st} 2016. This causes an increase in the number of
\emph{received} AIS messages (\emph{i.e.} points), and therefore an
apparent increase in the number of vessels and fishing hours. However,
the addition of new satellites affects all vessels in the same way.

Our analysis focuses on purse seine vessels, the most important fishery
for PNA
countries.\footnote{We perform some of the same analyses for longliners and include them in the Appendix}
We identify a total of 103 purse seiners that fished in PNA waters at
least once before
2015\footnote{New vessels that entered the fishery after 2015 were not exposed to the policy intervention in the pre-treatment period and are therefore excluded from our analyses.}.
These vessels represent over 26 million individual observations for the
2012 - 2017 period. We identify 65 vessels that have fished inside PIPA
at least once since 2012. From these, 62 did so before the announcement
(\emph{i.e.} 09/01/2014 \emph{sensu} \citep{mcdermott_2018}) but we are
only able to track 61 for the complete period of study before and after
its
implementation.\footnote{1 vessel never fished again after August 18, 2013.}
On the other hand, 38 vessels never fished inside PIPA but we only have
data for 26 of these vessels both before and after MPA implementation.

Therefore, our treatment group contains all purse seiners (n = 61) that
fished within PIPA at least once before the announcement, and that
continued to fish elsewhere after the January 2015 implementation.
Vessels in the control group meet the following two conditions: i)
never fished within PIPA waters from 2012-2015, and ii) vessels have fished in
surrounding areas (\emph{i.e.} PNA-countries' EEZ) before and after PIPA
closure (n = 26).

We include three additional control groups as a
robustness check. The first group contains only vessels that belong to PNA countries (n
= 7). The second group excludes Chinese vessels (n = 21). Our third control
is made up of Japanese purse seiners that fish in the Pacific but have
never fished inside PIPA (n =
27).\footnote{We also considered a control group of Taiwanese flagged vessels but there are only 4 vessels with pre-2015 data.}
Our main definition of treatment and control groups leaves us with 61 treated and 26
control vessels, which have just over 22 million
observations with about 22\% of these observations identified as fishing events by the neural network classification.

For each vessel, we calculate total daily fishing hours and build a daily panel
with 37,800 observations. Table \ref{tab:baci_n_s} shows the number
of vessels following a Before-After-Control-Impact (BACI) design, as well as the fishing hours, before
and after PIPA. Fig. \ref{fig:all_panels} shows that mean fishing hours
for purse seiners have an abrupt increase, just before January 1st,
2015. This trend is observed for both treated and control groups. Across
all measures, the treatment and control vessels follow similar patterns,
confirming our claim that the control group provides a plausible
counterfactual. Figure \ref{fig:baci_strict} in the Appendix provides a visual
representation of the vessel-level fishing events that make up each
group through time. We use this panel data to answer our key research questions: 
Are the 61 purse seiners that used to fish within PIPA affected by the closure? If so, how?
What are they doing differently and where are they going that they didn't go to before?
The following section describes our empirical identification strategy.

@@

@@

Our first analysis focuses on identifying the response of 
vessels to the PIPA closure. We use daily fishing hours, daily proportion of fishing vs.~non-fishing
hours, daily distance traveled (km), distance from shore (km) and
distance from home
port(km) as our main outcomes of interest.
We compare these outcomes before and after the implementation
of PIPA using a Difference-in-Differences approach. Our main
specification is the following:

\begin{figure*}[h]
\begin{align*}
log(y)_{i,t} = \alpha + \beta_1 Post_t + \beta_2 Treat_i + \beta_3 Post_t \times Treat_i + \phi_t + \gamma_i + \theta_t + \epsilon_{i,t}
\label{eqn:did}
\end{align*}
\end{figure*}

where \(y_{i,t}\) is the outcome of interest for vessel \(i\) on day \(t\). A dummy variable \(Post_t\) takes the value of 0 for all
dates prior to PIPA implementation and a value of 1 for all dates
following PIPA implementation. \(Treat_i\) is a dummy
variable indicating whether a vessel belongs to the treatment (\(Treat_i = 1\)) or control
(\(Treat_i = 0\)) group. \(\alpha\) is
the standard intercept term, \(\beta_1\) captures the temporal trend,
\(\beta_2\) captures the initial difference between treated and control groups,
and \(\beta_3\) is our parameter of interest: the Difference-in-Differences estimate capturing
the treatment effect. Finally, \(\phi_t\) and \(\gamma_i\) represent
month and flag dummies that account for seasonality or
country-level management interventions. \(\theta_t\) are time dummies to
account for the addition of more satellites over time.\footnote{We test for other more complex specifications that interact a quarterly dummy or year-month dummies with the treatment group and find qualitatively the same results.}

Our second part of the analysis focuses on the redistribution of fishing
effort. The role of institutions may play an important role. As stated
before, LSMPAs often span the entirety of an EEZ. In the case of purse
seining in the PNA, vessels purchase access to the fishery at the
beginning of the season (QUESTION FOR JC: WHEN IS THIS? And wasn't the closure expected?). If a vessel decides to fish within the PNA,
they've already made the decision to pay for access, expecting higher
returns from fishing there than in the high seas. In the particular case
of PIPA, one would expect that a vessel previously holding a permit to fish in
Kiribati waters would 
most likely continue to purchase access to PNA waters (either in Kiribati's remaining open areas or a nearby EEZ). Alternatively, if a vessel fishes illegally
in Kiribati waters before the implementation, one would expect them to
reallocate to other regions with equal or less enforcement. In this
case, they might then choose to move to other Kiribati waters, waters
of countries that have similar enforcement levels, or to the high seas.

We discretize spatial units by creating a polygon for
PIPA and distinct
spatial units for each geographically separate EEZ of each country. Some vessels might shift
from EEZs into the high seas, but we are interested in knowing
\emph{where} in the high seas, so we incorporate additional regions by
using a 1 degree buffer of the high seas around each of the EEZ regions.
The rest of the high seas are merged into a single spatial unit. For
example, if we were to do this only for Kiribati, we would have 8
spatial units: PIPA, three EEZ units, three 1-degree\footnote{Currently running the 5-degree version too} buffers of high seas
around each EEZ, and the rest of the high seas. We clipped any overlapping
high seas buffers to avoid duplication.

We then take these spatial polygons and rasterize them to a 1-degree grid
(see Figure \ref{fig:raster_rgn} in the Appendix), which we then use to rasterize points
of fishing activity. For each cell, we calculate the total monthly
fishing hours by treated and control vessels (see Figure
\ref{fig:fishing_raster}). Since we are interested in the redistribution
of fishing effort, we use this gridded data to calculate the proportion
of fishing hours that each region receives each month relative to the
total fishing hours observed across all regions and all cells in that
month.

@@

Our variable of interest, $y_{i,t}$ represents the proportion of
fishing hours that region $i$ receives at time $t$. Years are
modeled as factors ($Year_t$), using 2012 as the reference level.
\(Region\) is a dummy variable for regions defined above. Our vector of parameters of
of interest is $\beta_{3,i}$, which captures the yearly by-country
change in proportional allocation of fishing effort relative to 2012.

Finally, we inspect the crowding effects that may arise due to the reduction
in total fishing area. We use the rasterized fishing effort and calculate two
potential indices of crowding. For the first one, we count the number of cells
that had fishing activity from both groups over a given month. For the second one,
we calculate the correlation of presence/absence of fishing for both groups
over one month. If the implementation of the reserve induces a crowding effect,
we would expect to observe bo clear trend before the implementation of the reserve.
The implementation would be followed by an increase in these measures, which, 
if vessels avoid crowding, should then come back down. We therefore expect to
have three inflection points: 1) when there is an initial increase due to MPA
implementation, 2) When the crowding has reached its maximum and then decreases,
and 3) when the decrease in the measure levels off, pressumably to pre-MPA levels.
For this reason, we fit a 4th degree polynomial to our monthly indices. We do so by
centering our time series of crowding indices on the day of implementation. Our
explanatory variable is then the number of months before or after the implementation.
For example, since PIPA was implemented in January 1st of 2015, December of 2014 
has a value of -1, whicle Feb of 2015 would receive a value of 1.

@@

Regression coefficients for our DiD analysis are shown in Table
\ref{tab:did}. Columns 1 -3 use our primary control group. Columns 4 - 6 use
only PNA vessels as controls, and columns 7 - 9 exclude Chinese vessels.
Columns 10 and 11 use Japanese vessels in the Pacific as controls, and
don't include flag fixed-effects to avoid perfect collinearity
between the Japanese flag and
control group status.
Our DiD analysis shows an overall increase in purse seine fishing hours,
even after accounting for the introduction of new satellites (Table
\ref{tab:did}).\footnote{Results of the same analysis is shown for longliners in Table \ref{tab:long}} This 
\emph{post} coefficient estimate is consistent across different
model specifications and across controls, and
effectively represents the patterns observed in Figure
\ref{fig:all_panels}.

The \(\beta_3\) coefficient indicating our treatment effect suggests
that, relative to the control, treated vessels fish less: to the order
of 0.5-3.7 hours per day. In 6 out of 11 specifications, there is a statistically significant decrease in daily fishing hours. Another way to interpret this is that the increase
in fishing effort by treated vessels has occurred at a lower rate than
control vessels. These values are equivalent to a reduction of 15 - 111 fishing hours per month. However, this result is only robust and
significant (\(p < 0.01\)) in the simplest specification of the
Diff-in-Diff. When reducing the linear structure of the
\(Pre \times Post\) design and instead interacting the treatment dummy
with quarterly or year-month combinations, we are not able to reject the
null hypothesis of no change, suggesting that the total fishing hours
remains similar for the displaced vessels (Figs.
\ref{fig:q1}-\ref{fig:ym4}). Since vessels don't seem to be fishing a lot less, an obvious question arises: where are they
fishing now?

Figure \ref{fig:fishing_raster} provides a spatial representation of
these changes. Note that the 2013 effort distribution is roughly similar
across vessels. Then, in 2014, there is a sharp increase in fishing hours
by treated vessels inside PIPA (the so-called Blue Paradox). In 2015, treated vessels
then allocate more effort to the easternmost Kiribati EEZ and the high seas. In 2016, the spatial distribution of
fishing effort is again similar across groups. It is evident that the
increase in relative fishing effort is greater for regions closer to
PIPA.

Along a similar line, Figure \ref{fig:redist_trend} shows a detailed
temporal evolution of the relative allocation of fishing
effort.\footnote{Recall that to evaluate the redistribution of fishing effort we only track fishing vessels that belong to the treated group.}
It shows the change in fishing effort inside PIPA, including the
preemptive fishing and immediate reduction previously reported
\citep{mcdermott_2018}. Kiribati waters in KIR 1 and KIR 2 show an
increase in fishing effort for 2015; the same pattern is observed for
their respective High Seas buffers, and the general High Seas. Other
regions farther away from PIPA don't show a clear pattern under visual
inspection. This suggests fishing effort inside PIPA has moved to the high seas around Kiribati instead of paying to fish within alternative Kiribati EEZ waters.

We then move on to our regression results, presented as a figure showing
the year-region interaction coefficient estimates for each region as a
stream of changes through time (Figure \ref{fig:mean_change} and Table
\ref{tab:mean_change}). As noted before, the marginal change in the
relative allocation of fishing effort relative to 2012 increases in 2014
for PIPA. Most coefficients are not statistically significant, but
follow the same trends previously discussed, with a tendency for regions
closer to PIPA to show an increase in the relative allocation of fishing
effort in the post-implementation period. The largest increase is
observed for the Kiribati EEZ 2, where the redistribution of treated
vessels caused a 15\% increase in the \emph{relative} allocation of
fishing effort within Kiribati waters. This displacement is likely
causing a crowding effect after PIPA-fishing vessels redistribute to
other areas.

We use the rasterized effort to evaluate if the PIPA closure increased
crowding outside the LSMPA. We count the number of raster cells that
had both treated and control vessels each month (Fig.
\ref{fig:sp_corr}A). The pattern observed here could just be an artifact
of satellite detections increasing though time. Therefore we also
calculate a spatial correlation of presence/absence and observe similar
patterns (Fig. \ref{fig:sp_corr}B). Spatial overlap increases
in 2015 as vessels from PIPA are excluded, but then starts to decline.

\begin{figure}
\centering
\includegraphics{img/all_panels.pdf}
\caption{\label{fig:all_panels}Time seri showing mean monthly values es for our 10 variables of interest:
A) Fishing hours, B)Non-fishing hours at-sea,
C) Proportion of fishing hours to total hours at-sea,
D) Distance traveled, E) Mean distance from port,
F) Mean distance from shore,
G) Mean distance from port for fishing events,
H) Mean distance from shore for fishing events,
I) Proportion of hours spent in Kiribati waters,
J) Proportion of fishing hours spent in PNA waters.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{img/hist_kir_fishing.pdf}
\caption{\label{fig:hist_kir_fishing}Ridgeplot for the density of the \% of total fishing hours that take place
within Kiribati EEZ waters by year for treated vessels where the unit of observation is an individual vessel
}
\end{figure}

@@

%\clearpage
%
%\hypertarget{appendix}{%
%\section{Appendix}\label{appendix}}
%
%\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}
%
%\hypertarget{stream-of-data-by-vessel-group-and-period}{%
%\subsection{Stream of data by vessel, group, and
%period}\label{stream-of-data-by-vessel-group-and-period}}

@@

%\hypertarget{effort-redistribution}{%
%\subsection{Effort redistribution}\label{effort-redistribution}}
%\hypertarget{rasterized-regions}{%
%\subsubsection{Rasterized regions}\label{rasterized-regions}}
%\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-31-1.pdf}
%\caption{\label{fig:unnamed-chunk-31}\label{fig:raster_rgn}Rasterized
%regions. Each color (number) indicates a distinct region.}
%\hypertarget{fishing-raster}{%
%\subsubsection{Fishing raster}\label{fishing-raster}}
%\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-32-1.pdf}
%\caption{\label{fig:unnamed-chunk-32}\label{fig:fishing_raster_full}Yearly
%rasters of fishing hours for each group.}

>>>>

\begin{abstract}
Large-scale Marine Protected Areas (LSMPAs) have seen a significant
increase over the last years. The benefits of
reducing effort have been largely studied, but little empirical work has
evaluated how vessels react and redistribute after an MPA is created.
We use identification of fishing activity via
Automatic Identification Systems (AIS) and causal inference techniques
to provide the first analysis of behavioral changes and spatial
redistribution of tuna purse seiners due to the implementation of a
Large Scale Marine Protected Area in the Pacific Ocean. Our work
provides three main findings: 1) aggregate fishing effort remains
relatively unaffected; 2) vessels that fished inside the protected area
redistribute to adjacent waters; and 3) we observe a crowding effect for
the first months after implementation. Our results not only provide an
impact evaluation of the effect of LSMPAs on fishing activity, but
provide insights into vessel redistribution dynamics, which may have
ecological and economic implications for marine conservation and fisheries
management. As countries continue to implement LSMPAs as a way to
reach the stated 10\% target of ocean protection, managers should
consider how fishing effort will change in space and through time to
ensure that fishing effort is not just displaced elsewhere, leading to
overfishing in adjacent waters. While LSMPAs can provide a wide
range of benefits, their implementation must be accompained with
traditional fisheries management to maximize their effectiveness.

@@

\usepackage{lscape}
\begin{document}

@@

(QUESTION FOR JC: WHEN IS THIS (selling of rVDS rights)? And wasn't the closure expected?).

\section{Introduction}\label{introduction}

\dropcap{M}arine Protected Areas (MPAs) are intended to safeguard parts of the
ocean from fishing and other extractive activities. Current
international goals aim to protect 10\% of the ocean environment by
2020. In an effort to meet this target, there has been a rapid
increase recently in MPA coverage \citep{wood_2008,sala_2018}, largely driven by
a small number of Large Scale Marine Protected Areas (LSMPAs)\footnote{See, for example, \citep{game_2009,singleton_2014,boonzaier_2016,mccauley_2016,alger_2017}.}.
Today, a small number of LSMPAs make up at least 80\%
of the managed areas in the ocean \citep{toonen_2013}.

@@

\subsection{Large Scale Marine Protected Areas}\label{lsmpas}

@@

The amount of data gathered by GFW is dependent on the number of antennas and
satellites that can receive signals. The total satellite count increased from 3 to 6 on 
June 1\textsuperscript{st}
2014 , and then from 6 to 10 on January
1\textsuperscript{st} 2016. This causes an increase in the number of
\emph{received} AIS messages (\emph{i.e.} points), and therefore an
apparent increase in the number of vessels.
The addition of new satellites affects all vessels in the same way.

Our analysis focuses on tune purse seine vessels, the most important fishery
for PNA
countries.\footnote{We perform some of the same analyses for longliners and include them in the Appendix}
We identify a total of 186 purse seiners that fished in PNA waters at
least once before 2015. Vessels that entered the fishery after 2015 were not
exposed to the policy intervention in the pre-treatment period and are
therefore excluded from our analyses, leaving us with 125 vessels.
We identify 90 vessels that have fished inside PIPA
at least once since 2012; 72 did so before the announcement
of the closure (\emph{i.e.} 09/01/2014 \emph{sensu} \citep{mcdermott_2018}).
From these, we only observe before and after data for 64 tuna purse seiners.
On the other hand, 45 vessels never fished inside PIPA during the 2012 - 2015 period, 
but we only observe befora and after data for 28 of these.

Therefore, our treatment group contains all purse seiners (n = 64) that
fished within PIPA at least once before the announcement, and that
continued to fish elsewhere after the January 2015 implementation.
Vessels in the control group meet the following two conditions: i)
never fished within PIPA waters from 2012-2015, and ii) vessels have fished in
surrounding areas (\emph{i.e.} PNA-countries' EEZ) before and after PIPA
closure (n = 28). Together, these vessels represent more than 20 million georeferenced positions
for which we know activity (fishing or not fishing).

We include three additional control groups as a
robustness check. The first group contains only vessels that belong to PNA countries (n
= 7). The second group excludes all Chinese vessels (n = 12). Our third control
is made up of Japanese purse seiners that fish in the Pacific but have
never fished inside PIPA (n = 27). Our main definition of treatment and control
groups leaves us with 64 treated and 28 control vessels, which have just over 22 million
observations with about 22\% of these observations identified as fishing events by the neural network classification.

The behavioral responses that vessels can have to a spatial closure may occurr in different ways.
For example, displacement to new fishing grounds may represent a cost, as fishers search the ocean
to identify the most suitable fishing places. This may result in increased fuel and labor costs. Therefore,
we are interested in three key measures that could capture a captain's response to a spatial closure:
daily fishing hours, daily non-fishing at-sea hours, and the proportion of fishing to non-fishing hours at sea.
We calculate these measures for each vessel, and build a daily panel
with 152,802 observations. Table \ref{tab:baci_n_s} shows the number
of vessels following a Before-After-Control-Impact (BACI) design, as well as the mean fishing hours, before
and after PIPA was implemented.

Following the idea behind increased search time, we also use each vessel's track to compute
the total daily distance traveled (km) by calculating the distance between every observation.
Additionally, we include the daily mean distance to the nearest port and mean distance to shore
for all positions and for fishing events only. Finally, we calculate the monthly proportion of each vessel's
fishing time that was allocated to Kiribati waters and PNA waters

Fig. \ref{fig:all_panels} shows that monthly mean values of the 10 variables that we use,
for treated and control vessels.  Across all measures, the treatment and control vessels follow similar patterns,
confirming our claim that the control group provides a plausible
counterfactual.

@@

\begin{figure}
\centering
\includegraphics{img/all_panels.pdf}
\caption{\label{fig:all_panels}Time series showing mean monthly values es for our 10 variables of interest:
A) Fishing hours, B)Non-fishing hours at-sea,
C) Proportion of fishing hours to total hours at-sea,
D) Distance traveled, E) Mean distance from port,
F) Mean distance from shore,
G) Mean distance from port for fishing events,
H) Mean distance from shore for fishing events,
I) Proportion of hours spent in Kiribati waters,
J) Proportion of fishing hours spent in PNA waters.}
\end{figure}

@@

Our first analysis focuses on identifying the response of 
vessels to the PIPA closure. We use daily fishing and non-fishing hours,
daily proportion of fishing vs. non-fishing hours, daily distance traveled (km), 
distance from shore (km) and distance from home port(km) for fishing and non-fishing events,
and proportion of total fishing hours allocated to Kiribati waters and PNA waters
as our main outcomes of interest.
We compare these outcomes before and after the implementation
of PIPA using a Difference-in-Differences approach. Our main
specification is the following:

\begin{figure*}[h]
\begin{align*}
log(y)_{i,t} = \alpha + \beta_1 Post_t + \beta_2 Treat_i + \beta_3 Post_t \times Treat_i + \phi_t + \gamma_i + \epsilon_{i,t}
\label{eqn:did}
\end{align*}
\end{figure*}

where $log(y_{i,t})$ is the log-transformed outcome of interest for vessel $i$ on day $t$.
A dummy variable $Post_t$ takes the value of 0 for all
dates prior to PIPA implementation and a value of 1 for all dates
following PIPA implementation. $Treat_i$ is a dummy
variable indicating whether a vessel belongs to the treatment ($Treat_i = 1$) or control
($Treat_i = 0$) group. $\alpha$ is
the standard intercept term, $beta_1$ captures the temporal trend,
$beta_2$ captures the initial difference between treated and control groups,
and $beta_3$ is our parameter of interest: the Difference-in-Differences estimate capturing
the treatment effect. Finally, $phi_t$ and $\gamma_i$ represent
month and flag dummies that account for seasonality or
country-level management interventions. We test for other more complex specifications
that interact a quarterly dummy or year-month dummies with the treatment group
and find qualitatively the same results\footnote{I actually need more time to run these, but I don't think they'll change}.

Our second part of the analysis focuses on the redistribution of fishing
effort, where institutions may play an important role. As stated
before, LSMPAs often span the entirety of an EEZ. In the case of purse
seining in the PNA, vessels purchase access to the fishery at the
beginning of the season. If a vessel decides to fish within the PNA,
they've already made the decision to pay for access, expecting higher
returns from fishing there than in the high seas. In the particular case
of PIPA, one would expect that a vessel previously holding a permit to fish in
Kiribati waters would 
most likely continue to purchase access to PNA waters (either in Kiribati's
remaining open areas or a nearby EEZ). Alternatively, if a vessel fishes illegally
in Kiribati waters before the implementation, one would expect them to
reallocate to other regions with equal or less enforcement. In this
case, they might then choose to move to other Kiribati waters, waters
of countries that have similar enforcement levels, or to the high seas.

We discretize spatial units by creating a polygon for
PIPA and distinct
spatial units for each geographically separate EEZ of each country. Some vessels might shift
from EEZs into the high seas, but we are interested in knowing
\emph{where} in the high seas, so we incorporate additional regions by
using a 1-degree buffer of the high seas\footnote{Currently running the 5-degree version too}
The rest of the high seas are merged into a single spatial unit. For
example, if we were to do this only for Kiribati, we would have 8
spatial units: PIPA, three EEZ units, three 1-degree buffers of high seas  around each of the EEZ regions.
around each EEZ, and the rest of the high seas. We clipped any overlapping
high seas buffers to avoid duplication.

We then take these spatial polygons and rasterize them to a 1-degree grid
(see Figure \ref{fig:raster_rgn} in the Appendix), which we then use to rasterize points
of fishing activity. For each cell, we calculate the total monthly
fishing hours by treated and control vessels (Fig. \ref{fig:fishing_raster}).
Since we are interested in the redistribution
of fishing effort, we use this gridded data to calculate the proportion
of fishing hours that each region receives each month relative to the
total fishing hours observed across all regions and all cells in that
month.

@@

Our variable of interest, $y_{i,t}$ represents the proportion of
fishing hours that region $i$ receives at time $t$. Years are
modeled as factors ($Year_t$), using 2012 as the reference level.
$Region$ is a dummy variable for regions defined above. Our vector of parameters of
of interest is $\beta_{3,i}$, which captures the yearly by-country
change in proportional allocation of fishing effort relative to 2012.

Finally, we inspect the crowding effects that may arise due to the net reduction
in fishing area. We use the rasterized fishing effort and calculate two
potential indices of crowding. For the first one, we count the number of cells
that had fishing activity from treated and control vessels over a given month. For the second one,
we calculate the correlation of presence/absence of fishing events between both groups
over one month. If the implementation of the reserve induces a crowding effect,
we would expect to observe no trend before the implementation of the reserve.
The implementation would lead to an increase in these measures, which, 
if vessels avoid crowding, should then come back down. We therefore expect to
have three inflection points: 1) in the transition between no trend and initial crowding due to MPA
implementation, 2) When the crowding has reached its maximum and starts to decrease,
and 3) when the decrease levels off, pressumably to pre-MPA levels.
For this reason, we fit a 4th degree polynomial to our monthly indices. We do so by
centering our time series of crowding indices on the day of implementation. Our
explanatory variable is therefore the number of months before or after the implementation.
For example, since PIPA was implemented in January 1st of 2015, December of 2014 
has a value of -1, and Feb of 2015 would receive a value of 1.

@@

Regression coefficients for our DiD analysis are shown in Table
\ref{tab:main_DID}. Our DiD analysis shows that PIPA closure
had no effect on the number of fishing hours, but that it caused
non-fishing at-sea hours to increase by 1\%, This coefficient is
only marginally significant. The proportion of
fishing to non-fishing hours shows a positive change, but it is not
significant. However, treated vessels travel on average 19\% less
than control vessels, and opperate closer to shore and port.
Furthermore, PIPA implementation caused fishing events of treated
vessel to occur 23\% and 4\% closer to port and to shore, respectively.
Other specificationf that exclude fixed effects produce similar estimates (Fig. \ref{fig:other_specifications})
Since vessels don't seem to be fishing a lot less, an obvious question arises: where are they
fishing now?

%When reducing the linear structure of the
%\(Pre \times Post\) design and instead interacting the treatment dummy
%with quarterly or year-month combinations, we are not able to reject the
%null hypothesis of no change, suggesting that the total fishing hours
%remains similar for the displaced vessels (Figs.
%\ref{fig:q1}-\ref{fig:ym4}).

Figure \ref{fig:fishing_raster} provides a spatial representation of
these changes. Note that the 2013 effort distribution is roughly similar
across vessels. Then, in 2014, there is a sharp increase in fishing hours
by treated vessels inside PIPA. In 2015, treated vessels
then allocate more effort to the easternmost Kiribati EEZ and the high seas.
In 2016, the spatial distribution of
fishing effort is again similar across groups. It is evident that the
increase in relative fishing effort is greater for regions closer to
PIPA.

Along a similar line, Figure \ref{fig:redist_trend} shows a detailed
temporal evolution of the relative allocation of fishing
effort.  It shows the change in fishing effort inside PIPA, including the
preemptive fishing and immediate reduction previously reported
\citep{mcdermott_2018}. Recall that to evaluate the redistribution of fishing effort
we only track fishing vessels that belong to the treated group.
Kiribati waters show an increase in fishing effort for 2015;
the same pattern is observed for their respective High Seas buffers,
and the general High Seas. This suggests fishing effort inside PIPA has moved to the
high seas around Kiribati instead of paying to fish within alternative Kiribati EEZ waters.

We then move on to our regression results, presented as a figure showing
the year-region interaction coefficient estimates for each region as a
stream of changes through time (Fig. \ref{fig:mean_change} and Table
\ref{tab:mean_change}). As noted before, the marginal change in the
relative allocation of fishing effort relative to 2012 increases in 2014
for PIPA, followed by an abrupt reduction in 2015 and afterwards. This 
reduction confirms that effort within PIPA has been virtually eliminated.
Most coefficients are not statistically significant, but
follow the same trends previously discussed, with a tendency for regions
closer to PIPA to show an increase in the relative allocation of fishing
effort in the post-implementation period. The largest increase is
observed for the Kiribati EEZ 2 (Kiribati's easternmost EEZ), where the
redistribution of treated vessels caused a 10\% increase in the \emph{relative} allocation of
fishing effort within Kiribati waters. The high seas around Kiribati's EEZ exhibit
similar increases the year of implementation or the year after. Overall, the high seas
receive an incresse in the order of 5\% for 2015.This displacement is likely
causing a crowding effect after PIPA-fishing vessels redistribute to
other areas.

We use the rasterized effort to evaluate if the PIPA closure increased
crowding outside the LSMPA. We count the number of raster cells that
had both treated and control vessels each month (Fig.
\ref{fig:sp_corr}A). The pattern observed here could just be an artifact
of satellite detections increasing though time. Therefore we also
calculate a spatial correlation of presence/absence and observe similar
patterns (Fig. \ref{fig:sp_corr}B). Spatial overlap increases
in 2015 as vessels from PIPA are excluded, but then starts to decline.
Our 4th degree polynomial shows the expected patterh of a relative
constant trend before implementation which is followed by rapid increased
and subsequent decrease.

@@

\begin{figure}
\centering
	\includegraphics{img/hist_kir_fishing.pdf}
	\caption{\label{fig:hist_kir_fishing}Ridgeplot for the density of the \% of total fishing hours that take place
				within Kiribati EEZ waters by year for treated vessels where the unit of observation is an individual vessel.
}
\end{figure}

@@

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

\newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}

\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}

%
%\hypertarget{stream-of-data-by-vessel-group-and-period}{%
%\subsection{Stream of data by vessel, group, and
%period}\label{stream-of-data-by-vessel-group-and-period}}

@@

\hypertarget{effort-redistribution}{%
\subsection{Effort redistribution}\label{effort-redistribution}}

\hypertarget{rasterized-regions}{%
\subsubsection{Rasterized regions}\label{rasterized-regions}}

\begin{figure}
\centering
\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-31-1.pdf}
\caption{\label{fig:raster_rgn}Rasterized regions. Each color (number) indicates a distinct region.}
\end{figure}
%
%\hypertarget{fishing-raster}{%
%\subsubsection{Fishing raster}\label{fishing-raster}}
%
%\begin{figure}
%\centering
%\includegraphics{Manuscript_files/figure-latex/unnamed-chunk-32-1.pdf}
%\caption{\label{fig:unnamed-chunk-32}\label{fig:fishing_raster_full}Yearly
%rasters of fishing hours for each group.}
%\end{figure}
%
%\clearpage
